---
title: "cooredinated_ordering2"
author: "Diego Uchendu"
date: "30/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coordinated Ordering

1. Deterministic/continous Demand: The same amount of demand is there every instance of time and demand are met at every instance of time, thus no time varying demand

2. Instantaneous replenishment: As soon as you place an order,  the item arrives. Lead time= 0

3. No shortage: As soon as stock level on Y axis reaches 0, place an order for another quantity q. thus instanstaneous replenishment. 

In the data set "Data_ordering.xlsx" parts are represented by material ID column.
$i=$denotes part
$i \in \mathcal I \qquad \{i_1,i_2,...,i_{62}\}, \space$
$d_i=$ demand per day for part $i, \space$
$c_i^{or}=$ ordering cost for part $i ,\space$

$c_i^{sh}=$ stock holding cost rate based on unit price $p_i$ and interest rate $h, \qquad$.
$c_i^{sh}=p_i \cdot h$

calculate EOQ for each part i:
$$ q_i^* = \sqrt \frac {2 \cdot d_i \cdot (c_i^{or}+c^{-or})} {p_i \cdot h}$$
$\frac {d_i}{q_i}=$ no of orders per day for part $i, \qquad$
$\frac {q_i}{2}=$ average inventory for part $i$
$$ min \rightarrow (c_i^{-or}+ \sum_{i=1}^{n=62} \frac {d_i}{q_i} \cdot c_i^{or}) + (h \cdot \sum_{i=1}^{n=62} \frac {q_i}{2}\cdot p_i ) $$


```{r}
library(kableExtra)
library(knitr)
df <- data.frame(Cat = c("box 1i", "box 2i"), 
                 part_j1 = linebreak(c("yes", "no")), 
                 part_j2 = linebreak(c("yes", "no")),
                 part_j3=linebreak(c("no","yes")),
                 part_jn= linebreak(c("no", "yes")))
kable(df, col.names = c("boxID vs MatID", "part_j1", "part_j2","part_j3","part_jn"), escape = F, caption = "parts each boxes can contain") %>%
  kable_styling(latex_options = "hold_position")
```
$b_{i}=$capacity of a box for part $i$ given in pieces
$b_{j}=$ Size of box $j$
$$b_{j}= box_{length} \times box_{width} \times box_{height}$$
Assuming $q_i=50$ for part $i$ and $b_i=45$ then number of boxes ordered is $\frac {q_i}{b_i}= \frac {50}{45}=1.1=1$ because only complete boxes can be ordered

# Optimum Ordering constraint

$$x_i=\frac {q_i}{b_i}=Number \space boxes \space ordered  \\ 
\frac {d_i}{b_i}=Number \space boxes \space demanded \\
T= \frac {q_i}{d_i}=TimeCycle$$   
Constraint:
$$\frac {q_i}{b_i} - \frac {d_i}{b_i} \cdot T \le 0$$
# Rack Length or one runway constraint

$n_k=$Number of boxes in a runway $k$ with box ID $j$ for part $i\\$:
E.g: box ID 3103147 has a width of 198 but Rack length is 6000,
maximum number of box 3103147 in runway is $n_k= \frac {6000}{198}=30.30=30 \space boxes\\$

$$ n_{k}=\frac {rack_{length}}{b_{ji\space length}} \\ $$
$x_{i}=$ lets say that there are 720 boxes with id 3103147 containing part i Mat ID(7346592+72), the first 30 boxes are assigned to runway $k_1 \qquad$ 

$x_{ik}=$ number of boxes ordered for part $i$ in runway $k, \qquad$ $b_{ji\space length}$ is the length of box id $j$ containing part $i,\space \\$
$x_{ik}= \frac {q_{ik}}{b_{ik}} \\$


$$\sum_{i \in \mathcal I } x_{ik} \cdot b_{ji\space length} \le n_{k} \qquad \forall k \in \mathcal K $$
# Rack Width

Collapsing Rack 4 levels in a rack and joining the 8 racks to become 1 level.

Total rack width available: 
$rack_{total_{Width}}=rack_{width} \times 4 \times 8\\$
Sum of number of parts $i$ is equal to box size for part $i$.

$$\sum_{k \in \mathcal K } x_{kj} \cdot b_{ji\space width} \le rack_{total_{Width}}$$




```{r}

```


```{r pressure, echo=FALSE}

```

