---
title: "coordinated_ordering_3"
author: "Group B"
date: "08/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Constraints

$b_{sorting}=$ sorting column lenght or width.This determings the rack width

$b_{-sorting}=$if sorted by length then the value is width and vice versa, 

E.g Using box_ID 6203060, sorted by length
$b_{sorting}=396$, $b_{-sorting}=297$
Also for material ID 7305667+74 with $q_i=214$ therefore,
$rack_{length}=6000$
1. Fraction of $q_i$ in a rack length:
$$q_i \cdot \frac {b_{-sorting}}{rack_{length}}=214 \cdot \frac {297}{6000}=10.593= 10 boxes $$
10 boxes for material ID 7305667+74 in a lane or rackway.

(2) How many lanes part $i$ will occupy if you order some number of boxes.
$$\frac {2 \cdot  q_i \cdot b{-sorting}}{rack_{length}}= \frac {2 \cdot 214 \cdot 297}{6000}=21.186=22 lanes$$
This box needs 22 lanes or runways.

(3) Fraction of Runway width occupied by box i : 
$$b_{sorting}/rack_{totalWidth}$$

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
