---
title: "coordinated_ordering_3"
author: "Group B"
date: "08/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coordinated Ordering

1. Deterministic/continous Demand: The same amount of demand is there every instance of time and demand are met at every instance of time, thus no time varying demand

2. Instantaneous replenishment: As soon as you place an order,  the item arrives. Lead time= 0

3. No shortage: As soon as stock level on Y axis reaches 0, place an order for another quantity q. thus instanstaneous replenishment. 

In the data set "Data_ordering.xlsx" parts are represented by material ID column.
$i=$denotes part
$i \in \mathcal I \qquad \{i_1,i_2,...,i_{62}\} \\$
$d_i=$ demand per day for part $i \\$

# data Extraction

```{r}
library("readxl")
product_data <- read_excel("Data_ordering.xlsx",sheet = "product data")
box_data <- read_excel("Data_ordering.xlsx",sheet = "box data")

#rack data
Total_racks=8
levels_per_rack=4
rack_length= 6000
rack_width=1750
rack_height=300

```

# Convert to Boxes
Convert demand $d_i$, quantity $q_i$ and unit price$p_i$ all from part $i$ to box $bi$ which is the capacity of a box for part $i$ 

$de_i=\frac {d_i}{b_i} \rightarrow$ Demand for box with part $i \\$
$\mathcal pr_i= b_i \cdot p_i \rightarrow$ Price for box with material $i \\$
$de_i = demand\_per\_year \\ ,$
$pr_i = box\_cost$

```{r}
#assuming there are 365 working days in a year
#converting everything to boxes

product_data$demand_per_year= ceiling((product_data$`demand per day` *365)/ product_data$`pieces/box`)

product_data$box_cost= product_data$`pieces/box` * product_data$price

product_data
```

Append Ordering cost for each item

```{r}

order_cost <- double(length(product_data$`box ID`))

  for(j in 1: length(box_data$`box ID`)){
    for (k in 1:length(product_data$`box ID`)) {
      if(box_data$`box ID`[j]==product_data$`box ID`[k]){
        order_cost[k] <- box_data$`ordering cost (€)`[j]
      }
    }
    
  }

product_data$ordering_cost <- order_cost 
product_data
```
$c_i^{or}=$ ordering cost for part $i ,\space$

$c_i^{sh}=$ stock holding cost rate based on unit price $p_i$ and interest rate $h, \qquad$.
$c_i^{sh}=p_i \cdot h$

## Separate Ordering (SO)
calculate EOQ for each part i:
$$ q_i^* = \sqrt \frac {2 \cdot de_i \cdot (c_i^{or}+c^{-or})} {pr_i \cdot h}$$
A. Calculate The EOQ:
```{r}
#EOQ
#dei= demand_per_year, cori= ordering cost for box i
#cord= ordering cost, pri= box cost, h= interest rate
so_eoq_fun <- function(dei,cori,cord,pri,h){
  eoq<- sqrt((2*dei*(cori+cord))/(pri*h))
  
  return(eoq)
}

vec_so_eoq_fun <- Vectorize(so_eoq_fun(dei=product_data$demand_per_year,cori=product_data$ordering_cost,cord=1500,pri=product_data$box_cost,h=0.10))

#Append Vector eoq to the table
product_data$eoq <- ceiling(vec_so_eoq_fun)
```


B Number of lanes you occupy with EOQ

## Constraints

$b_{i(sorting)}=$ sorting column lenght or width.This determines the rack width

$b_{i(-sorting)}=$if sorted by length then the value is width and vice versa, 

Add $b_{i(sorting)}$ and $b_{i(-sorting)}$ to product_data table
```{r}
#create a space holder
b_sorting <- double(length(product_data$`box ID`)) 
b_not_sorting <-double(length(product_data$`box ID`))

  for(j in 1: length(box_data$`box ID`)){
    for (k in 1:length(product_data$`box ID`)) {
      if(box_data$`box ID`[j]==product_data$`box ID`[k]){
        
        if(box_data$sorting[j]=="width"){
          b_sorting[k] <- box_data$width[j]
          b_not_sorting[k]<- box_data$length[j]
          
        }else{
          b_sorting[k] <- box_data$length[j]
          b_not_sorting[k]<- box_data$width[j]
        }
      }
    }
    
  }


product_data$b_sorting <- b_sorting 
product_data$b_not_sorting <- b_not_sorting
```


E.g Using box_ID 6203060, sorted by length
$b_{i(sorting)}=396$, $b_{i(-sorting)}=297$
Also for material ID 7305667+74 with $q_i=214$ therefore,
$rack_{length}=6000 \\$
1. Number of boxes in a lane:
$$n_i=\frac {rack_{length}}{b_{i(-sorting)}}=\frac {6000}{297}=20.20=20boxes$$





2. How many lanes part $i$ will occupy if you order some number of boxes
$$lane_i=\frac {q_i}{n_i}\\lane_i=q_i \cdot \frac {b_{i(-sorting)}}{rack_{length}}=214 \cdot \frac {297}{6000}=10.593= 11 \space lanes $$

```{r}
lane <- ceiling(product_data$eoq * (product_data$b_not_sorting/rack_length))
lane
```

(C) Do we meet the capacity contraint?

3. Total number of lanes constraints

Collapsing Rack 4 levels in a rack and joining the 8 racks to become 1 level.

Total rack width available: 
$$rack_{total_{Width}}=(rack_{width} \times 4 \times 8)$$
```{r}
rack_total_width <- rack_width * 4 * 8
rack_total_width #56000 mm
```


Please note the coefficients are  are different for every ith item except racklength which is the same for all items.Also the left and right hand side of the equation need to be in mm.
$$\sum_{i=1}^{n=62}lane_i \cdot b_{i(sorting)}\le rack_{Total_{width}} \\ \sum_{i=1}^{n=62} q_i \cdot \frac {b_{i(-sorting)} \cdot b_{i(sorting)}}{rack_{length}} \le rack_{Total_{width}} \quad \quad (1)$$

```{r}
#get the with of the lanes in mm
rack_width_occupied=sum(sum(lane*product_data$b_sorting))
rack_width_occupied
if(rack_width_occupied <= rack_total_width){
  print(past0("Capacity constraint fulfilled: ", rack_width_occupied, "<=",rack_total_width))
 
}else {
  violated <- rack_width_occupied - rack_total_width
  print(paste0("Capacity constraint violated by: ",violated ))
}

```
As seen above contraint was voilated by $1387664 mm$, this means that we ordered too much, therefore we need to optimize $q_i$


#Prove equation 1 for  single item using same figures above:
Using our example to get number of the total width i.e (summation of lanes) in $mm$: 
Summation of lanes can simply be: 
$$lanes_i \cdot b_{i(sorting)}= 10.593 \cdot 396=4194.828$$
Using equation (1) to prove this concept.

$$214 \cdot \frac {396 \cdot 297}{6000}=4194.828 \\ b_{isorting}= \frac {4194.828}{10.593}=396$$

(D) Adjust Q by reducing it:

$\frac {d_i}{q_i}=$ no of orders for part $i \\$
$\frac {q_i}{2}=$ average inventory for part $i$



$$ min \rightarrow (\sum_{i=1}^{n=62}c^{-or} + \sum_{i=1}^{n=62} \frac {de_i}{q_i} \cdot c_i^{or}) + (h \cdot \sum_{i=1}^{n=62} \frac {q_i}{2}\cdot pr_i ) $$
Subject to:
$$\sum_{i=1}^{n=62} q_i \cdot \frac {b_{i(-sorting)} \cdot b_{i(sorting)}}{rack_{length}} \le rack_{Total_{width}} $$

Calculating using the current EOQ
```{r}

#dei=product_data$demand_per_year,cori=product_data$ordering_cost,cord=1500,pri=product_data$box_cost,h=0.10

so_obj_funct<-function(cord,cori,dei,eoq,h,pri){
  
  obj <- (62*cord)+(sum((dei/eoq)*cori))+ (h*sum((eoq/2)*pri))
  return(obj)
}

so_obj_funct(1500,product_data$ordering_cost,product_data$demand_per_year,product_data$eoq,0.10,product_data$box_cost)

```


# Optimization with respect to capacity contraint
